\Large \bf Polyglots : Crossing Origins by Crossing Formats

### 2013-02-21 main.tex
Major updates of mitigtion and some in Vulnerabilities and attacks

Other potential targets for this scenario are job broker services and conference 
submission systems, which by design allow submission of PDFs to be viewed by
users.

@@

\subsection{Payloads}

The consequences of exploiting an infiltrated origin depend on the capabilities 
of the format used. These consequences span from forging requests to the vulnerable website,
abusing the credentials of the victim, to extracting sensitive information about the user that is
stored on the vulnerable website.

@@

\paragraph{Cross-origin leakage}

@@

\begin{lstlisting}[label=lst:embed-pdf, caption=HTML for embedding PDF content]
<embed type="application/pdf" 
src="vulnerable.com/content?input=%PDF...">
\end{lstlisting}

@@

\subsection{Content smuggling}
\todo{Deal with this section, extend and polish}
Due to the nature of the PDF, it can without much effort be combined with just 
about any other format. This provides ample opportunity to create malicious 
polyglots where PDF is either the benign or the malicious format. Consequently, 
this significantly expands the attack surface, making it important to take 
measures to protect against these attacks.

\paragraph{PDF as the benign format}
\todo{Services that accept PDFs, PDFs can be used to hide another malicious format}
Services like job brokers commonly let the user upload a CV in the form of a PDF-file.
Before such PDF-files are published to recruiters, they are verified to not 
contain any malicious payload. An attacker can produce a PDF that is valid and 
benign, but also a polyglot hiding another malicious format, such as Flash. 
A proof-of-concept can easily be created by storing the PDF as a static string 
variable in the malicious flash source code. When the flash is compiled, the 
source of the PDF is part of the output, making it a valid PDF.

\paragraph{PDF as the malicious format}
 Given the extensive capabilities of PDF, exploiting a content smuggling 
vulnerability with a PDF-based polyglot attack, can be done without much effort.
A server-side verification process will base its verification on the benign format,
and is likely to miss the malicious PDF components. To prevent the attack, the 
verification process have to actively search for and remove PDF specific syntax.

\subsection{PDF communication methods}
%\todo{rewrite section, focus on embedded flash, xxe and code samples}
%
In order to extract information and leak across origins, a communication channel
is required. PDF documents provide multiple ways of generating HTTP requests. 
Many of these are cross-origin, but, out of security concerns, only one-way in the sense that 
the PDF document will never see the result. Therefore, the focus is directed towards 
two methods that allow bidirectional communication. Since the document will retain the 
origin from which it was served, all 
requests issued from the document will include any cookies associated with the target origin.

@@

\begin{lstlisting}[label=lst:xml-pdf, caption=PDF using XML for communication]
%PDF-
1 0 obj<<>>stream
xml = '<!DOCTYPE foo[<!ENTITY x SYSTEM "'+
	URL+'">]><foo>&x;</foo>'
var doc = XMLData.parse(xml);
getURL('attacker.com/?'+doc.saveXML())
endstream
trailer<</Root
   <</Pages<<>>
      /OpenAction
      <</S/JavaScript/JS 1 0 R>>
   >>
>>
\end{lstlisting}

@@

The conclusion is 
that nine websites out of the hundred apply insufficient content filtering with 
respect to the PDF format. Out of the nine found to be vulnerable;three were susceptible to traditional 
XSS and thereby also to PDF polyglot attacks, two were susceptible only to PDF based polyglot attacks, and four applied 
insufficient content filtering, but the input was reflected in a way that prevented exploitation.
%
%, which have the potential of compromising the 
%confidentiality and integrity of the website's users.
%
We are in contact with the teams responsible for the websites to help
mitigate the vulnerabilities.

@@

%1-2 pages 
\section{Mitigation}
\label{sec:mit}
This section gives advice on various mitigation approaches available for each of the
affected components, from server to client.

@@

As a content provider on the Internet today there are precautions that one can 
take to mitigate this class of vulnerabilities. Which precautions to take depend 
on the kind of services provided. The mitigation recommendations for syntax 
injection apply to services that rely on user input, and the content smuggling
recommendations apply to services that serve user-supplied files.

@@

To prevent PDF-based syntax injection attack, the task is simpler. The token-set identified in 
Section~\ref{sec:pdf-injection}, are essential 
to create valid PDF syntax, namely \emph{'$<$', '$>$', and '/'}. Filtering 
user input to remove or encode these characters effectively mitigates the 
vulnerability.

@@

The current best-practice recommendation from Google~\cite{Google:ContentHosting} 
on hosting user-supplied content, is to serve the content from an origin that is 
completely separate, and isolated from the sensitive services. It also describes how 
Google moved away from sanitizing content, as it was found to be insufficient in 
many cases. These best-practices successfully prevents content-smuggling 
attacks in all aspects, as the infiltrated origin is separate from the origin 
containing sensitive data.

\todo{What about serving as attachment, so that it is not viewed within the browser? How does it work with embedded content?}
\todo{Problem of transferring credentials between domains for services where the content requires authentication}

@@

In the case of PDF-based polyglots, the browser can intervene when there is a 
mismatch between the content-type provided by the server and the type attribute 
of the \emph{object} tag. By acknowledging that there can be a potential security issue in 
interpreting the supplied content in the requested format and alerting the user to this
threat, the vulnerability can be mitigated.

@@

As a general rule of thumb, the interpreter at the very least alert the user if the served 
content-type differs from the expected. A preferred alternative is to 
not attempt to interpret the content at all. This holds true especially when 
the served content-type is well known and radically different from that which 
the interpreter is designed for.

As for the PDF file format, the underlying design decisions have led to the 
current parsing being very relaxed, even with respect to the specification. 
Making the parsing more strict and enforcing many of the specified requirements 
will reduce the attack surface. It is, however, not sufficient for complete 
mitigation because the PDF format is a container format it is designed to 
embed syntax from other files. Even when parsing strictly according to 
the specification, it is a simple task to create a PDF-based polyglot.

@@

We also plan to extend the empirical study to popular services like
HotCRP, Europass CV, and arXiv.org that allow uploading PDF
documents. Authorization is pending for an in-depth security study of
a major conference submission system.

>>>>

%Other potential targets for this scenario are job broker services and conference 
%management systems, which by design allow submission of PDFs to be viewed by
%users.

@@

\subsection{Payloads}
\label{sec:payloads}
The consequences of exploiting an infiltrated origin depend on the capabilities 
of the format used. These consequences span from forging requests to the vulnerable website,
abusing the credentials of the victim, to extracting sensitive information about the user that is
stored on the vulnerable website.

@@

\paragraph{Cross-origin information leakage}

@@

\begin{lstlisting}[label=lst:embed-pdf, caption=HTML for embedding PDF content]
<embed type="application/pdf" 
src="vulnerable.com/?input=%PDF...">
\end{lstlisting}

@@

\subsection{Content smuggling}

Due to the nature of the PDF, it can without much effort be combined with just 
about any other format. This provides ample opportunity to create malicious 
polyglots where PDF is either the benign or the malicious format. Consequently, 
this significantly expands the attack surface, making it important to take 
measures to protect against these attacks.

\paragraph{PDF as the benign format}
%\todo{Services that accept PDFs, PDFs can be used to hide another malicious format}
Services like job brokers commonly let the user upload a CV in the form of a PDF-file.
Before such PDF-files are published to recruiters, they are verified to not 
contain any malicious payload. Such a verification only extends to the PDF format itself.
An attacker can produce a PDF that is valid and 
benign, but also a polyglot hiding another malicious format, such as Flash. As 
described in the content smuggling scenario in Section~\ref{sec:smuggling}, the 
uploaded PDF file is then embedded on the attacker's web site, but now as the 
malicious format. Using social engineering, the attacker can persuade 
the victim to visit the web site.

Creating a PDF/Flash polyglot is no major challenge.
A proof-of-concept can easily be created by storing the PDF source code as a static string 
variable in the malicious Flash source code. When compiling Flash, the output is 
compressed by default to save space, thereby obfuscating the PDF source code. However, tools exist 
that decompress the Flash files, which restores the plain PDF code.

\paragraph{PDF as the malicious format}
As detailed in the content smuggling scenario in Section~\ref{sec:smuggling}, the 
attacker uploads the PDF polyglot to a vulnerable content hosting service. 
A server-side verification process will base its verification on the benign formats 
it expects to receive. The polyglot is designed to verify correctly as the benign format,
and the verification is likely to miss the malicious PDF components, as it is unaware of the alternate interpretation of the content.

Given the extensive capabilities of the PDF format, exploiting a content smuggling 
vulnerability with a PDF-based polyglot attack, can be done without much effort.
To prevent the attack, the 
verification process have to actively search for and remove PDF specific syntax.

\subsection{PDF payloads}
%\todo{rewrite section, focus on embedded flash, xxe and code samples}
%
As discussed in Section~\ref{sec:payloads}, there are two approaches to exploiting 
these vulnerabilities; cross-site request forgery, and cross-origin information leakage.
Both of these require extracting information from the vulnerable service, which is something that can be achieved using the capabilities of the PDF format.

In order to extract information and leak across origins, a communication channel
is established. PDF documents provide multiple ways of generating HTTP requests; 
many of which allow cross-origin communication, but are, out of security concerns, only one-way in the sense that 
the PDF document will never see the result. Therefore, the focus is directed towards 
two methods that allow bidirectional communication; XML external entities, and embedded Flash. 
 Since the document will retain the 
origin from which it was served, all 
requests issued from the document will include any cookies associated with the target origin.

@@

\begin{lstlisting}[label=lst:xml-pdf, caption=PDF using XML for communication]
%PDF-
1 0 obj<<>>stream
xml = '<!DOCTYPE foo[<!ENTITY x ' +
 'SYSTEM "'+URL+'">]><foo>&x;</foo>';
var doc = XMLData.parse(xml);
getURL('attacker.com/?secret='+
  doc.saveXML())
endstream
trailer<</Root
   <</Pages<<>>
      /OpenAction
      <</S/JavaScript/JS 1 0 R>>
   >>
>>
\end{lstlisting}

@@

The conclusion is 
that nine websites out of the hundred apply insufficient content filtering with 
respect to the PDF format. Out of the nine found to be vulnerable;three were susceptible to traditional 
XSS and thereby also to PDF polyglot attacks, two were susceptible only to PDF based polyglot attacks, and four applied 
insufficient content filtering, but the input was reflected in a way that prevented exploitation, e.g. the header appeared after 1024 bytes.
%
%, which have the potential of compromising the 
%confidentiality and integrity of the website's users.
%
We are in contact with the teams responsible for the websites to help
mitigate the vulnerabilities.

@@

%1-2 pages 
\section{Mitigation}
\label{sec:mit}
This section gives advice on various mitigation approaches available for each of the
affected components, both server-side and client-side. It should be noted that 
preventing polyglot attacks in the general case is a complex task, as one would 
need to take all potentially malicious formats in to account. Mitigating instances of 
polyglot attacks based on particular formats are significantly more straightforward.

@@

As a content provider on the Internet today there are precautions that one can 
take to mitigate this class of vulnerabilities. Which precautions to take depend 
on the kind of services provided. The mitigation recommendations for syntax 
injection apply to all services that rely on user input, and the content smuggling
recommendations apply to services that serve user-supplied files.

@@

To prevent PDF-based syntax injection attack, the task is simpler. As discussed in Section~\ref{sec:pdf-injection}, the token-set identified in 
Section~\ref{sec:pdf-injection}, are essential 
to create valid PDF syntax. %, namely \emph{'$<$', '$>$', and '/'}.
Filtering 
user input to remove or encode these characters effectively mitigates the 
vulnerability. 
Luckily, because of the significant overlap with the token-set
of HTML, many of the contexts where user input can occur are already being protected.
Special attention is required in contexts are not traditionally filtered for 
HTML tokens, e.g. JSON.

@@

The current best-practice recommendation from Google~\cite{Google:ContentHosting} 
on hosting user-supplied content, is to serve the content from a sandboxed origin that is 
completely separate, as per SOP, and isolated from the sensitive services. 
These best-practices successfully prevents content smuggling 
attacks, as the restrictions of the SOP prevents the content from accessing any sensitive 
resources in the origin of the web service.

These recommendations comes with a caveat that need to be taken into consideration. Some user supplied 
content is only meant to be accessible to certain authenticated users, e.g. photos that 
is only shared with friends. In that case, the service needs to transfer the 
credentials required to authenticate the user from the sensitive origin to the sandboxed 
origin, without actually revealing the credentials used in the sensitive origin. 
Revealing the credentials, e.g. using the same cookies on both origins, defeats 
the purpose of the sandbox origin entirely. Currently there is no uniform solution 
to this problem. Some common solutions are encountered: using the hashed 
credentials of the sensitive origin; or generating public, but obscured, links 
that are later shared manually.

%\todo{Problem of transferring credentials between domains for services where the content requires authentication}

@@

In the case of PDF-based polyglots, the browser can intervene when there is a 
mismatch between the content-type provided by the server and the type attribute 
of the \emph{object} tag. The vulnerability can be mitigated by acknowledging that there can be a potential security issue in 
interpreting the supplied content in the requested format and alerting the user to this
threat.

@@

As a general rule of thumb, the interpreter must at the very least alert the user if the served 
content-type differs from the expected. A preferred alternative is to 
not attempt to interpret the content at all. This holds true especially when 
the served content-type is well known and radically different from that which 
the interpreter is designed for.

As for the PDF file format, the underlying design decisions have led to the 
current parsing being very relaxed, even with respect to the specification. 
Making the parsing more strict and enforcing many of the specified requirements 
will reduce the attack surface. However, it is not sufficient for complete 
mitigation. As discussed in Section~\ref{sec:pdf-based-polyglots} the PDF format is a container format; designed to 
embed syntax from other files. Even when parsing strictly according to 
the specification, it is a simple task to create a PDF-based polyglot.

@@

We also plan to extend the empirical study to popular services like
HotCRP, Europass CV, and arXiv.org that allow uploading PDF
documents. Authorization is pending for an in-depth security study of
a major conference management system.
